---
title: "Week 2 Exercises"
author: "Nick Hegge"
date: "July 11, 2024"
output: pdf_document
---

Please complete all exercises below. You may use stringr, lubridate, or the forcats library. 

Place this at the top of your script:
library(stringr)
library(lubridate)
library(forcats)

# Exercise 1
Read the sales_pipe.txt file into an R data frame as sales. 

```{r}

library(stringr)
library(lubridate)
library(forcats)

data <- read.csv("C:/Users/Nhegg/OneDrive/Documents/Database/Merrimack College - Nick Masters/DSE5002/Module_1_fork/DSE5002_Module2/sales_pipe2.csv",
                 stringsAsFactors = FALSE, fileEncoding = "latin1")



```

# Exercise 2
You can extract a vector of columns names from a data frame using the colnames() function. Notice the first column has some odd characters. Change the column name for the FIRST column in the sales date frame to Row.ID. 

**Note: You will need to assign the first element of colnames to a single character.**

```{r}
colnames(data)[1] <- "Row.ID"


```

# Exercise 3
Convert both Ship.Date and Order.Date to date vectors within the sales data frame. What is the number of days between the most recent order and the oldest order?
How many years is that?
How many weeks?

**Note: Use lubridate**

```{r}

# Convert Ship.Date and Order.Date to date vectors
data$Ship.Date <- as_date(data$Ship.Date, format = "%m/%d/%Y")
data$Order.Date <- as_date(data$Order.Date, format = "%m/%d/%Y")

# Calculate number of days between the most recent order and the oldest order
days_diff <- as.numeric(difftime(max(data$Order.Date), min(data$Order.Date), units = "days"))

# Calculate number of years and weeks
years_diff <- days_diff / 365.25
weeks_diff <- days_diff / 7

# Print
cat("Number of days between the most recent order and the oldest order:", days_diff, "days\n")
cat("Number of years:", years_diff, "years\n")
cat("Number of weeks:", weeks_diff, "weeks\n")



```


# Exercise 4
What is the average number of days it takes to ship an order?

```{r}
# Average
average_days_to_ship <- mean(data$Ship.Date - data$Order.Date, na.rm = TRUE)

# Print
cat("Average number of days it takes to ship an order:", round(average_days_to_ship, 2), "days\n")


```


# Exercise 5
How many customers have the first name Bill?
You will need to split the customer name into first and last name segments and then use a regular expression to match the first name bill. Use the length() function to determine the number of customers with the first name Bill in the sales data. 

```{r}
# Split customer name into first and last name
data$Customer.FirstName <- word(data$Customer.Name, 1)

# Find 'Bill'
bill_customers <- str_detect(data$Customer.FirstName, regex("^Bill$", ignore_case = TRUE))

# Count customers with the first name Bill
number_of_bills <- length(which(bill_customers))

# Print
cat("Number of customers with the first name Bill:", number_of_bills, "\n")



```

# Exercise 6
How many mentions of the word 'table' are there in the Product.Name column? 
**Note you can do this in one line of code**

```{r}
# Count 'table' in the Product.Name column
table_mentions <- sum(str_count(data$Product.Name, regex("table", ignore_case = TRUE)))

# Print
cat("Number of mentions of the word 'table' in the Product.Name column:", table_mentions, "\n")



```

# Exercise 7
Create a table of counts for each state in the sales data. The counts table should be ordered alphabetically from A to Z. 

```{r}
# table of counts for each state
state_counts <- table(data$State)

# Convert to order alphabetically
state_counts_df <- as.data.frame(state_counts)
state_counts_df <- state_counts_df[order(state_counts_df$Var1), ]

# Print
print(state_counts_df)



```

# Exercise 8
Create an alphabetically ordered barplot for each sales Category in the State of Texas. 

```{r}
library(ggplot2)

# Filter State of Texas
texas_data <- subset(data, State == "Texas")

# Create the barplot
ggplot(texas_data, aes(x = fct_inorder(Category))) +
  geom_bar() +
  labs(title = "Sales Categories in Texas",
       x = "Category",
       y = "Count") +
  theme_minimal() +
  coord_flip()




```

# Exercise 9
Find the average profit by region.
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
# Calculate the average profit by region
average_profit_by_region <- aggregate(Profit ~ Region, data = data, FUN = mean, na.rm = TRUE)

# Print
print(average_profit_by_region)



```

# Exercise 10
Find the average profit by order year. 
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
# Extract the order year from Order.Date
data$Order.Year <- year(data$Order.Date)

# Average
average_profit_by_year <- aggregate(Profit ~ Order.Year, data = data, FUN = mean, na.rm = TRUE)

# Print
print(average_profit_by_year)


```

